Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-05-24 doc/bcspaper/article.tex
Resized figure 1 to make more room and make Latex position it more suitably

However, the document title is just one part of the important structure to be
found within a scientific document.  Liakata \emph{et al.} (2012) describe a
system for automatically processing and classifying sentences in a research
paper according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent academic
information retrieval system can use this information in order to provide
better filtering and search capabilities for researchers.  Partridge {\st the
ability to search for phrases and keywords by CoreSC will facilitate}
implements such context-aware keyword search, by allowing researchers to search
for papers where a term appears in sentences with a specific CoreSC label (e.g
only in {\em Method} senetences). This can be used to greatly improve both the
precision with which researchers are able to perform searches for scientific
literature and the accuracy of those searches in terms of relevance to the
reader.

@@

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes.  The PLoS
journals\footnote{PLoS: \url{http://www.plos.org/}} contain large volumes of OA
literature under a permissive license that allows data mining.  They also use
the Pubmed Central markup schema, which is compatible with SAPIENTA, for papers
published through their journal.  The PLoS journals advanced search offers
approximately 50 types of paper through which to restrict the search.  Many of
these paper type categories contain too few papers to be useful, others are too
specific (e.g.``Message from PLoS'') or appear rather adhoc.  Indeed it is not
clear how these paper types have been identified. % in conclusions add how this
can give recommendations for Plos and other journals We chose to look at a
range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Synopsis,
Perspective, Viewpoint, Opinion, Review, Research).

The papers were downloaded from PLoS using {\em plosget.py}, a short python
script that uses the PLoS restful search API. The number of papers downloaded
per paper type category was as follows: 200 Essay, 99 Correspondence, 174
Synopsis, 200 Perspective, 74 Viewpoint, 93 Opinion, 107 Case Study, 312
Review, 200 Research.  These formed a corpus of 1459 papers.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX \cite{PDFX}, a free service hosted
by the University of Manchester (\url{http://pdfx.cs.man.ac.uk/}).
PDFX also provides output that already has some metadata, such as
title, author, and abstract, associated with it.  SAPIENTA allocates
each sentence within a paper a separate CoreSC label.  However, the
sentence boundaries must be detected before the document is passed
into SAPIENTA for annotation.  The sentence splitter uses a regular
expression to split the sentences within the document and then adds
the necessary \verb|<s>| tags to the markup to indicate the location
of each sentence.  Once the paper has been split, SAPIENTA is used to
annotate each sentence with a relevant CoreSC label.  These labels are
assigned using a conditional random fields model based upon a number
of features such as the sentence's location within the paper and pairs
and triplets of words found consecutively within the sentence
\cite{Liakata2012}.  The SAPIENTA system has been trained on a corpus
of 265 chemistry and biochemistry papers and we have done no domain
adaptation for the papers in PLoS.

\begin{figure}[t]
\sidecaption[t]
% Use the relevant command for your figure-insertion program
% to insert the figure file.
% For example, with the option graphics use
%\includegraphics[scale=.65]{figure}
%
% If no graphics program available, insert a blank space i.e. use
%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
%
\centering
%\begin{subfigure}[b]{0.6\textwidth}
\includegraphics[width=\textwidth]{figures/review_corescs.png}
%\caption{CoreSC content for a sample review paper}
%\end{subfigure}
%\begin{subfigure}[b]{0.6\textwidth}
\includegraphics[width=\textwidth]{figures/report_corescs.png}
%\caption{CoreSC content for a sample research paper}
%\end{subfigure}
\caption{The CoreSC content of a review paper and a research paper randomly selected from the corpus}
%\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page. If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
\label{fig:coresc_pies} 
\end{figure}

>>>>

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent academic
information retrieval system can use this information in order to provide
better filtering and search capabilities for researchers.  Partridge {\st the
ability to search for phrases and keywords by CoreSC will facilitate}
implements such context-aware keyword search, by allowing researchers to search
for papers where a term appears in sentences with a specific CoreSC label (e.g
only in {\em Method} senetences). This can be used to greatly improve both the
precision with which researchers are able to perform searches for scientific
literature and the accuracy of those searches in terms of relevance to the
reader.

@@

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes.  The PLoS
journals\footnote{PLoS: \url{http://www.plos.org/}} contain large volumes of OA
literature under a permissive license that allows data mining.  They also use
the Pubmed Central markup schema, which is compatible with SAPIENTA, for papers
published through their journal.  The PLoS journals advanced search offers
approximately 50 types of paper through which to restrict the search.  Many of
these paper type categories contain too few papers to be useful, others are too
specific (e.g.``Message from PLoS'') or appear rather adhoc.  Indeed it is not
clear how these paper types have been identified. We chose to look at a range
of types, some of which we expect to overlap or have an unclear distinction.
These are namely: Essay, Correspondence, Case Study, Synopsis, Perspective,
Viewpoint, Opinion, Review, Research.

% in conclusions add how this can give recommendations for Plos and other journals 
The papers were downloaded from PLoS using {\em plosget.py}, a short python
script that uses the PLoS RESTful search API. The number of papers downloaded
per paper type category was as follows: 200 Essay, 99 Correspondence, 174
Synopsis, 200 Perspective, 74 Viewpoint, 93 Opinion, 107 Case Study, 312
Review, 200 Research.  These formed a corpus of 1459 papers.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted
by the University of Manchester (\url{http://pdfx.cs.man.ac.uk/}).
PDFX also provides output that already has some metadata, such as
title, author, and abstract, associated with it.  SAPIENTA allocates
each sentence within a paper a separate CoreSC label.  However, the
sentence boundaries must be detected before the document is passed
into SAPIENTA for annotation.  The sentence splitter uses a regular
expression to split the sentences within the document and then adds
the necessary \verb|<s>| tags to the markup to indicate the location
of each sentence.  Once the paper has been split, SAPIENTA is used to
annotate each sentence with a relevant CoreSC label.  These labels are
assigned using a conditional random fields model based upon a number
of features such as the sentence's location within the paper and pairs
and triplets of words found consecutively within the sentence
\cite{Liakata2012}.  The SAPIENTA system has been trained on a corpus
of 265 chemistry and biochemistry papers and we have done no domain
adaptation for the papers in PLoS.

\begin{figure}[t]

\centering

\includegraphics[width=0.75\textwidth]{figures/review_corescs.png}

\includegraphics[width=0.75\textwidth]{figures/report_corescs.png}

\caption{The CoreSC content of a review paper and a research paper randomly selected from the corpus}

\label{fig:coresc_pies} 
\end{figure}
